|

### 2020-06-30 paper.Rmd
Rewriting the parametrised rescaling section.

Eq. \ref{eq:resc} is fixed for a given value of $p$, but we may wish to tune the rescaling more freely. One way of doing this is to replace the dimensionality $p$ by the tuned $p_{eff} = s p$. When $s<1$ the rescaling will be softer, and $s>1$ results in more aggressive rescaling than suggested by $p$ alone. Note that when $p_{eff} < 2$ we actually invert the behaviour and shift the focus away from the center, in general this is not interesting. This is shown in Fig. \ref{fig:scaled} for $p=10$ and selected values of $s$. XXXX this is basically the same plot as before... is there a better illustration to show here, or maybe I don't need any?

```{r scaled, fig.cap="Relation between $r$ and $r'$ for $p=10$ and different values of the scaling $s$. We can tune $s$ to control how strong the rescaling is, with $s<1$ leading to a weaker effect, and large values of $s$ increase the effect.", fig.height=3, fig.width=6, out.width="60%", fig.align = "center"}
library(tidyverse)
# define index as function of c
trans_s <- function(s, p){
  p <- s * p
  function(r){
    sqrt( 1 - (1-r^2)^(p/2))
  }
}

>>>>

Eq. \ref{eq:resc} is fixed for a given input dataset, by evaluating the number of dimensions $p$ and the maximum distance from the center $R$. However, we may wish to tune the rescaling. For example, the dimension of the input might not reflect the intrinsic dimensionality of the dataset. This could be the case when dimension reduction was used prior to visualisation, e.g. displaying only the first few principal components. In this case the effective dimensionality $p_{eff}$ is likely between the original dimensionality and the selected number of principal components, since some directions would be considered pure noise, while others might still carry relevant information.

Similarly, the overall scale of the transformation is determined by $R$. In the case of an approximately spherical and uniform distribution the maximum distance from the center works well and ensures the validity of the rescaling in Eq. \ref{eq:resc}. But it is not robust and might result in a much larger scale than desired, especially in the case of outlying observations. Tuning the scale would also allow increased zooming near the center without changing $p_{eff}$.
XXX When selecting a value $R$ that is smaller than the maximum distance from the center, we need to ensure that the projected radius of points is always smaller than $R$ before rescaling according to Eq. \ref{eq:resc}. We do this by replacing $r_i$ as
\begin{equation}
r'_i = \min(r_i, R)
\end{equation}
for each observation $i = 1, ..., n$.

In practice we allow tuning $p_{eff} = s p$ by selecting the scaling parameter $s$. By default, $s=1$ and $p_{eff}=p$. When $s<1$ the rescaling will be softer, and $s>1$ results in more aggressive rescaling than suggested by $p$ alone. Note that when $p_{eff} < 2$ we actually invert the behaviour and shift the focus away from the center, in general this is not wanted.

On the other hand, $R$ can be directly supplied by the user, and is set to the maximum distance from the center if no value is supplied. XXX it is important to ensure that all variables are on a comparable scale before applying this method!

```{r scaled, eval=F, fig.cap="Relation between $r$ and $r'$ for $p=10$ and different values of the scaling $s$. We can tune $s$ to control how strong the rescaling is, with $s<1$ leading to a weaker effect, and large values of $s$ increase the effect.", fig.height=3, fig.width=6, out.width="60%", fig.align = "center"}
library(tidyverse)
# define index as function of c
trans_s <- function(s, p){
  p <- s * p
  function(r){
    sqrt( 1 - (1-r^2)^(p/2))
  }
}
